\newcommand\Lim[1]{{\tt Lim}(#1)}
\newcommand{\NFZ}{{\tt NF}_0}

\section{Solvable terms and Limits}

In this section we formally define the notion of limit of a reduction on a cyclic 
$\lambda$-term and we prove its main properties. The notion of limit is akin to the notion
of B\"{o}hm tree. However, there is some important difference with B\"{o}hm trees,
due to the fact that we are modeling equality of functions and functionals over $\N$,
instead of pure $\lambda$-terms as in the case of B\"{o}hm trees. 


\subsection{Limit reductions of LAMBDA}
     
Let $S$ be the
set of all first symbols of terms of $\LAMBDA$, namely: 
$x^T_i, \lambda x^T_i, \ap, 0, \Succ, \cond$, for all types $T$.

\begin{definition}[Finite paths of a term]
Let $n \in \N$ and $t,u \in \LAMBDA$.  
A finite path $\pi$ from $t$ to $u$, just a path for short, 
is any finite sequence $\pi:[0,n] \rightarrow \{1,2\} \times S$, such that for some
$t_0, \ldots, t_{n} \in \LAMBDA$ we have:

\begin{enumerate}
\item 
$t_0 = t$ and $t_{n} = u$
\item 
for all $i \in [0,n]$, if $\pi(i)=(j,s) \in \{1,2\} \times S$, then $s$ is the first symbol of $t_i$;
and if $i \ge 1$ then $t_{i}$ is the immediate subterm 
number $j$ of $t_{i-1}$, otherwise $j=1$ (a dummy value). 
\end{enumerate}
We denote with $\universe{t}$ the set of paths of $t$.
\end{definition}

A term $t \in \LAMBDA$ is uniquely defined from the set $\universe{t}$ of its paths:
if $\universe{t} = \universe{u}$ then $t = u$ as labeled trees.

An infinite reduction sequence from $t$ is any infinite sequence 
$\rho : \N \rightarrow \LAMBDA$ such that $t = \rho(0)$ and $\rho(i) \reduces \rho(i+1)$
for all $i \in \N$. A path $\pi$ is definitively in $\rho$ if and only if 
$\exists i \in \N. \forall j \in \N. (j \ge i \implies \pi \in \universe{\rho(j)}$.
We define a limit-reduction relation $t \reduces_\rho u$ if and only if
$\universe{u}$ is exactly the set of paths $\pi$ definitively in $\rho$. 

\begin{proposition}
For every $t \in \LAMBDA$ and any infinite reduction $\rho$ from $t$ there is a
unique $u \in \LAMBDA$ such that $t \reduces_\rho u$.
\end{proposition}

This is akin to the notion of infinite $\lambda$-term in Klop. 
A term can have more than one limit-reduction, because we can reduce infinitely many times a 
term while avoiding to contract one particular redex. However, as remarked by Klop, at least one 
limit is normal. Infinite Church-rosser implies that the normal limit is unique. 


\subsection{Solvable Terms and Limit Normal Forms of LAMBDA}
An alternative definition of Limit Normal Form for $\LAMBDA$
can be given by adapting the notion of solvable $\lambda$-term of finite $\lambda$-calculus.
This second definition is equivalent for terms satisfying GTC.

We define a main path $\pi$ of a term $t \in \LAMBDA$.
$\pi$ is the unique path of $t$ moving left except in any subterm $\cond(a,b)(u)$, 
where it moves right. We define a head redex of $t$ as the first redex in the main path $\pi$, 
if any. We say that $t$ is in head normal form if $\pi$ is finite and $\pi$ has no head redex. 
$t$ is solvable if for some $u \in \LAMBDA$ we have $t \reduces^* u$ 
and $u$ is in head normal form. 
$t$ is unsolvable if $t$ is not solvable. 

Some remark.The main path is non-progressing, because in $\cond(a,b)$ the path moves to $a$. 
Thus, if a term satisfies GTC then its main path $\pi$ is finite, otherwise it wo
and the term is head normal if and only if it has no head redex.
 
The limit normal form of $t$ is bottom (undefined) if $t$ is unsolvable, otherwise the limit is obtained by reducing $t$ to some head normal form, taking the first symbol of the head normal form and then taking the limit of all immediate subterms, cyclically. 

%Limit is not unique because we can reduce infinitely many times a term while avoiding to contract one particular redex. However, as remarked by Klop, at least one limit is normal. Infinite Church-rosser implies that the normal limit is unique. 
Having the same normal limit can be expressed by a maximal bisimulation 
(see later in this section).
 
      Without GTC, the limit normal form can be bottom, or can include some bottom subterms, as it happens for B\"{o}hm trees in $\lambda $-calculus.  In our paper, we can prove that all limit of all terms with GTC include no bottom. This is a corollary of our normalization result: head-reduction is 0-safe and therefore it always terminates in a head-normal form. GTC is preserved by reductions and subterms. If we recursively head-normalize immediate subterms of a head-normal form we always reduce terms with GTC, therefore we always obtain solvable subterms.
 
 
 
\subsection{A co-inductive Definition of Equality between normal limits of terms}

\vspace{10pt}

\begin{remark}
  We assume the following (consequence of limit Church-Rosser):
  \begin{center}
    Any $t\in \GTC$ has a unique normal form $\Lim{t}$ (limit of safe-reductions).
  \end{center}
\end{remark}

\noindent{\bf Purpose of this sub-section}\quad
We want to characterize the limit equivalence of terms in $\GTC$, 
which means two terms have the same limit normal form: $\Lim{t} = \Lim{u}$. 
More formally, we try to coinductively define a binary relation $\approx$ that satisfies:
\[
t \approx u \quad \Longleftrightarrow \quad \Lim{t} = \Lim{u}. 
\]

A {\em 1-context} is a finite context with several number of distinguished holes $[-]$.

\begin{definition}[0-context]
  \begin{center}
    $C ::= x \mid \lambda x.C \mid CC \mid 0 \mid \Suc{C} \mid \Cond{C}{[-]}$
  \end{center}
\end{definition}

We write $C[-,-,-]$ if $C$ has three holes, and write $C[t,u,v]$ for the resulting term
obtained by filling terms $t,u,v$ for the holes. 

A term can be divided into the 0-context part and the terms part filling the holes of the 0-context,
namely the following lemma holds. 
\begin{lemma}
  For any $t\in \LAMBDA$, there uniquely exists $(C,\vec{u})$ such that $t = C[\vec{u}]$. 
\end{lemma}

For a binary relation $R$ on terms,
we write $(t_1,\ldots,t_n)R(u_1,\ldots,u_n)$ if $t_iRu_i$ holds for all $i$. 

Let $\NFZ$ be the set of $0$-safe normal forms. 
Note that if both $C_1[\vec{t_1}]$ and $C_2[\vec{t_2}]$ are $0$-safe normal forms of $t$, then $C_1=C_2$.
  
\begin{definition}[0-bisimulation]\rm
  A binary relation $R$ on $\LAMBDA$ is said to be a {\em 0-bisimulation} if it satisfies the following:
  \begin{center}
    $tRu$ and $t \nsafeReducesAst{0} C_t[\vec{t}] \in \NFZ$ and $u \nsafeReducesAst{0} C_u[\vec{u}] \in \NFZ$
    \\
    \quad
    implies
    \quad
    $C_t = C_u$ and $\vec{t}R\vec{u}$. 
  \end{center}
  
  We define $\approx$ by the largest $0$-bisimulation.
\end{definition}

\begin{lemma}
  $\approx$ is an equivalence relation.
\end{lemma}
\begin{proof}
  Reflexivity for $\approx$ holds by $=\,\subseteq\,\approx$, since $=$ is a $0$-bisimulation.
  Symmetricty holds by $\approx^{-1}\,\subseteq\,\approx$, since $\approx^{-1}$ is a $0$-bisimulation.
  Transitivity holds by $(\approx\circ\approx)\,\subseteq\,\approx$, since $\approx\circ\approx$ is a $0$-bisimulation.
  \hfill$\Box$
\end{proof}

Finally we show the relation $\approx$ characterizes the limit equivalence. 
\begin{proposition}
  $t \approx u$ and $\Lim{t} = \Lim{u}$ are equivalent. 
\end{proposition}
\begin{proof}
  $(\Leftarrow)$: Let $R$ be the limit equivalence relation, namely $tRu \Leftrightarrow \Lim{t}=\Lim{u}$. 
  Then $R$ is a $0$-bisimulation.
  Hence $\Lim{t} = \Lim{u}$ implies $t \approx u$ by $R \subseteq \approx$. 
  
  $(\Rightarrow)$:
  We show this direction by proof by contradiction.
  Define
  \[
  X = \{(t,u) \mid \text{$t \approx u$ and $\Lim{t} \neq \Lim{u}$}\,\}.
  \]
  Assume that $X \neq \emptyset$ (there is a counter example of the $(\Rightarrow)$-direction). 
  For $(t,u)\in X$, we define the difference depth $\sharp(t,u)$
  by $\min\{\,|\pi| \mid (\Lim{t})_{\pi} \neq (\Lim{u})_{\pi}\,\}$. 
  Fix a $(t_0,u_0) \in X$ that has the least difference depth.
  Take $0$-safe normal forms $C_t[\vec{t'}]$ and $C_u[\vec{u'}]$ of $t_0$ and $u_0$, respectively. 
  Then, by $t_0\approx u_0$, we have $C_t = C_u$ and $\vec{t'} \approx \vec{u'}$. 
  Also we have $C_t[\overrightarrow{\Lim{t'}}] = \Lim{t_0}$ and $C_u[\overrightarrow{\Lim{u'}}] = \Lim{u_0}$. 
  By $\Lim{t_0}\neq \Lim{u_0}$ and $C_t=C_u$,
  there is $j$ such that $\Lim{t'_j} \neq \Lim{u'_j}$. 
  Hence we have $(t'_j,u'_j) \in X$. 
  However we also have $\sharp(t'_j,u'_j) < \sharp(t_0,u_0)$,
  since, when we consider the path to $t'_j$ from the root of $t_0$,
  there must be at least one ${\tt cond}$ on the path.
  This contradicts the leastness of $(t_0,u_0)$.
  Therefore we have $X = \emptyset$, namely the $(\Rightarrow)$-direction holds.
  \hfill$\Box$
\end{proof}


\subsection{Comparing with B\"{o}hm trees}
We only consider equational theories equating all unsolvable. We consider all unsolvable 
denoting a unique \quotationMarks{undefined} value.
      
In $\lambda$-calculus two solvable lambda terms can be equated in a consistent 
theory equating all unsolvable if and only if have the same  
$\eta$-bohm-tree, that is, have the same limit normal $\beta\eta$-form. 
 
Instead, we found no simple way to characterize the terms
we may consistently equate in $\LAMBDA$, or in $\CTlambda$ by adding a reduction
like $\eta$. 
We include some example, showing that this problem is difficult.

    
\begin{enumerate}

\item
An example of a set of terms of $\LAMBDA$ we can equate:
$$\lambda x.\Succ(x), 
\ \ \ 
\cond(1,\cond(2, \cond(3, \ldots))), 
\ \ \
\cond(\Succ(0),\lambda x.\Succ^2(x))
$$
All terms above are notations for the successor map, therefore they can be consistently equated. 

\item
As an opposite example, let $I = \lambda x.x$ and 
$f = \cond(I(0),f)$ and $g = \cond(I(0),\cond(1,g))$. 
$f$ and $g$ are non-terminating terms of $\CTlambda$: they include infinitely many
disjoint redexes of the form $I(0)$. They are solvable terms with limits 
representing different map, which cannot be equated. 
The limit normal form of $f$ is $\cond(0,\cond(0, \cond(0, ...)))$: the map $f$ is always $0$.
The limit normal form of $g$ is $\cond(0,\cond(1, \cond(0, ...)))$: the map $g$ return $0$ or $1$
according if the argument is even or odd. We cannot equate $f=g$,
otherwise we obtain $0 = f(1) = g(1) = 1$, and from it the equality of all terms.
\end{enumerate}
 
We could not find finitely many computable reduction rules such that two terms have the same
limit normal form if and only if they represent the same map on $\N$.
 
  
